Inferring disease architecture and predictive ability\\with LDpred2-auto

### 2023-08-31 paper/paper-infer-supp.tex
add dense fine-mapping

>>>>

\begin{figure}[p]
	\centerline{\includegraphics[width=\textwidth]{manhattan_height_one_region}}
	\caption{GWAS of height in a restricted region containing the most significant HapMap3+ variant for height. \label{manhattan_height}}
\end{figure}

\begin{figure}[p]
	\centerline{\includegraphics[width=\textwidth]{pip_height_one_region}}
	\caption{Posterior inclusion probabilities (PIPs) for the variants in Figure \ref{manhattan_height}. 
	L is the maximum number of causal variants in SuSiE-RSS, and max(p) is the maximum value of the estimated $p$ in LDpred2-auto.
	Both panels A and B represent the same data, but colors and facets are changed. \label{pip_height}}
\end{figure}

### 2023-08-31 paper/paper-infer.tex
add dense fine-mapping

@@

Our proposed method has limitations. First, when power is low (i.e.\ when $N h^2 / p$ is low), estimates of $\alpha$ and $p$ become less reliable.
Therefore, we recommend to use the 3-parameter model (with $\alpha$, when using \texttt{use\_MLE = TRUE}), but only when power is sufficient.
We also recommend using option \texttt{allow\_jump\_sign = FALSE} for robustness \cite[]{prive2021identifying}, and for getting accurate or conservative $p$ estimates. When still obtaining a large $p$ estimate ($>1\%$) with this option, we recommend rerunning LDpred2-auto without this option to get a less conservative estimate (cf.\ simulations).
However, LDpred2-auto estimates of $h^2$ and $r^2$ seem always reliable, except for height for which they are probably overestimated.
We think this is likely due to assortative mating, which causes GWAS effects to be inflated \cite[]{border2022assortative,herzig2023model}, which causes our estimates to be inflated as well. 
Second, the $h^2$ from LDpred2-auto is also slightly overestimated when using a small LD reference panel or when the reference panel does not closely match with the ancestry of the GWAS summary statistics.
Future work could focus on correcting these issues.
Third, when using external GWAS summary statistics, it is often beneficial to regularize the LD matrix (via parameter \texttt{coef\_shrink}, especially when using the new sampling and model), however it leads to an increased estimation for e.g.\ $r^2$ and $h^2$.
Future work could focus on correcting these estimates, and also on identifying the best shrinkage regularization coefficient to apply based on e.g.\ some distance metric (mismatch) between GWAS summary statistics and the LD reference used.
Fourth, because we use a limited set of variants as input for LDpred2, causal variants identified by LDpred2-auto are probably tagging variants that are highly correlated with unobserved causal variants close by. 
LDpred2-auto can also totally miss some causal variants when they are poorly tagged by the set of variants used.
Future work will focus on scaling LDpred2-auto to using more variants to alleviate this limitation; currently, the main issue is the large size of the LD matrix that grows quadratically with the number of variants used.

>>>>

Finally, since the HapMap3+ variants still represent a rather small proportion of common variants, we showcase running LDpred2-auto using a more dense set of variants from a small region, as usually done for fine-mapping.
We identify the most significant HapMap3+ variant for height, rs2871960 on chromosome 3, and read all the variants within a 500kb distance that have a MAF larger than 0.005 and INFO score larger than 0.5 in the UK Biobank; there are 3881 such variants.
Then, we perform a GWAS for these variants using the same training individuals as used before (Figure \ref{manhattan_height}).
Finally, using the resulting GWAS summary statistics and an LD reference from the same set of individuals, we run both SuSiE-RSS and LDpred2-auto (\_nojump). We test two different values for L, the maximum number of causal variants in SuSiE-RSS (10 and 100), and three different values for the maximum value of the estimated $p$ (1, 0.01, and 0.001).
Per-variant posterior probabilities of being causal (also known as posterior inclusion probabilities, PIPs) are very similar with both methods, especially when restricting $p$ in LDpred2-auto, which is similar to what SuSiE-RSS assumes (more conservative).

@@

Our proposed method has limitations. First, when power is low (i.e.\ when $N h^2 / p$ is low), estimates of $\alpha$ and $p$ become less reliable.
Therefore, we recommend to use the 3-parameter model (with $\alpha$, when using \texttt{use\_MLE = TRUE}), but only when power is sufficient.
We also recommend using option \texttt{allow\_jump\_sign = FALSE} for robustness \cite[]{prive2021identifying}, and for getting accurate or conservative $p$ estimates. When still obtaining a large $p$ estimate ($>1\%$) with this option, we recommend rerunning LDpred2-auto without this option to get a less conservative estimate (cf.\ simulations).
However, LDpred2-auto estimates of $h^2$ and $r^2$ seem always reliable, except for height for which they are probably overestimated.
We think this is likely due to assortative mating, which causes GWAS effects to be inflated \cite[]{border2022assortative,herzig2023model}, which causes our estimates to be inflated as well. 
Second, the $h^2$ from LDpred2-auto is also slightly overestimated when using a small LD reference panel or when the reference panel does not closely match with the ancestry of the GWAS summary statistics.
Future work could focus on correcting these issues.
Third, when using external GWAS summary statistics, it is often beneficial to regularize the LD matrix (via parameter \texttt{coef\_shrink}, especially when using the new sampling and model), however it leads to an increased estimation for e.g.\ $r^2$ and $h^2$.
Future work could focus on correcting these estimates, and also on identifying the best shrinkage regularization coefficient to apply based on e.g.\ some distance metric (mismatch) between GWAS summary statistics and the LD reference used.
Fourth, because we use a limited set of variants as input for LDpred2, causal variants identified by LDpred2-auto are probably tagging variants that are highly correlated with unobserved causal variants close by. 
LDpred2-auto can also totally miss some causal variants when they are poorly tagged by the set of variants used.
Future work will focus on scaling LDpred2-auto to using more variants to alleviate this limitation; currently, the main issue is the large size of the LD matrix that grows quadratically with the number of variants used.
For now, one can concentrate on small regions of the genome, where LDpred2-auto can be rerun using a much more dense set of variants, as typically done in fine-mapping analyses (cf.\ Results section ``Local heritability and polygenicity'').
